---
title: "CS 422"
author: "Julen Ferro Bañales. Master's in CDS&OR, Illinois Institute of Technology"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---

<!-- More information in R Markdown can be found at:
1. https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_RMarkdown.html  This is 
   the place to start since it is a short tutorial.
2. https://rmarkdown.rstudio.com/index.html This contains a longer 
   tutorial.  Take a look at the cheatsheet in 
   https://rmarkdown.rstudio.com/lesson-15.html, it is a concise 
   reference of R Markdown on two pages.
<-->

## Use this as a template for your homeworks.
#### Rename it to firstname-lastname.Rmd.
#### Run all the chunks by clicking on "Run" at the top right of the edit 
#### window and choose "Run All".  Assuming there were no errors in the
#### chunk, you should see a "Preview" button become visible on the top
#### left of the edit window.  Click this button and a html document should
#### pop up with the output from this R markdown script.

### Part 2.1.a)

Importing required modules.

```{r}

library(factoextra)
library(ggplot2)
library(ggpubr)


```

i) 

```{r}

set.seed(123) # setting a seed for the pseudo-random numbers
rm(list=ls())
df <- read.csv(file = "file19.txt",     # TXT data file indicated as string or full path to the file
           header = TRUE,    # Whether to display the header (TRUE) or not (FALSE)
           sep = "",          # Separator of the columns of the file
           dec = ".", 
           skip = 20, 
           strip.white = TRUE,
           skipNul = TRUE)         # Character used to separate decimals of the numbers in the file

plot(df, main="Raw points")
typeof(df)
head(df)


```

ii) 

```{r}



```

iii) 

```{r}



```


### Part 2.1.b)

i) 

```{r}
#Getting the optimal number of clusters

dataframe <- df[, 2:9]

fviz_nbclust(dataframe, kmeans, method = "wss") +
geom_vline(xintercept = 3, linetype = 2)


```

ii) 

```{r}

# Running K-means

k <- kmeans(dataframe, centers=3, nstart=25)
fviz_cluster(k, data=dataframe)


```

iii) 

```{r}

for i in range 1:3{
 print(paste0("In the ", k, "cluster, there are: ", n, "observations" ))
}



```

iv) 

```{r}



```

v) 

```{r}



```

vi) 

```{r}



```


### Part 2.2.a)


```{r}



```

### Part 2.1.b)

i) 

```{r}



```

ii) 

```{r}



```

### Part 2.1.c)

i) 

```{r}



```

ii) 

```{r}



```

iii) 

```{r}



```

### Part 2.1.d)

i) 

```{r}



```

ii) 

```{r}



```

iii) 

```{r}



```

### Part 2.1.e)

i) 

```{r}



```

ii) 

```{r}



```

### FINAL