---
  title: "Homework 7"
output: 
  html_notebook:
  toc: yes
toc_float: yes
---
  
```{r}

library(keras)
library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)

rm(list=ls())

# Set working directory as needed
setwd("...")

df <- read.csv("wifi_localization.csv")

# Seed the PRNG
set.seed(1122)
df <- df[sample(nrow(df)), ] # Shuffle, as all of the data in the .csv file
                             # is ordered by label!  

train_proportion <- 0.8
sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(train_proportion, 1-train_proportion))
df_train  <- df[sample, ]
df_test   <- df[!sample, ]

```
# --- Your code goes below ---

#DATA DESCRIPTION
#The CSV file contains 2001 rows, one of them is the header row
#Each row contains 8 columns (7 wifis and 1 room number)

# (a)


```{r}

model <- rpart(room ~., data = df_train, method = 'class')
rpart.plot(model, extra = 106)


```

# (b)
# Note that in (b) either use a new variable to store the model, or null out
# the variable that stored the model in (a) if you want to reuse that variable.
# The reason is that if you don't null it out, the model in (b) will have
# residual information left over from (a) and your results will not be quite
# accurate.


```{r}



```


# (c) 


```{r}



```


# (d)


```{r}



```
